LAGOS lake depth

### 2020-11-23 figures/01_geometry.Rmd
propagate bathy metric name changes, calculate mape model metric, closes #17

```{r slope-compare, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height=3.5}
# compare inlake slope, buffer slope, and online slope
# setwd("figures") 
bp <- readRDS("../data/00_bathy_depth/bathy_pnts.rds")
dt <- read.csv("../data/00_geometry/nearshore.csv", stringsAsFactors = FALSE) %>%
  mutate(llid = as.character(llid))
dt <- left_join(dt, st_drop_geometry(bp), 
                by = "llid") %>%
  dplyr::filter(!is.na(inlake_slope_mean))

# jsta::jheatmap
jheatmap(
  dplyr::select(dt, 
                slope_online_mean,
                inlake_slope_mean,
                slope_mean, 
                slope_mean_ne,
                dist_deepest, 
                dist_viscenter, 
                maxdepth,
                meandepth))

GGally::ggpairs(dplyr::select(dt, 
                inlake_slope_mean, inlake_slope.y,
                slope_online_mean,
                slope_mean, 
                # dist_deepest, 
                # dist_viscenter, 
                maxdepth), 
                progress = FALSE, 
                ) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1), 
          strip.text.y = element_text(angle = 0))

@@

# dt[22,]
# which(dt$inlake_slope.y > 0.4)
# which(dt$inlake_slope.x > 0.4)
dt <- dplyr::filter(dt, inlake_slope.y < 0.4)

cowplot::plot_grid(
  gg_scatter_slope("slope_mean", "maxdepth", dt),
  gg_scatter_slope("inlake_slope.y", "maxdepth", dt), 
  gg_scatter_slope("slope_mean", "inlake_slope.y", dt)
  )
```

```{r buffer-vs-inlake_slope, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.width=4}
lgus_pred <- read.csv("../data/lagosus_depth_predictors.csv", stringsAsFactors = FALSE) %>%
  mutate(lagoslakeid = as.character(lagoslakeid))
test <- left_join(dt, lgus_pred, by = c("llid" = "lagoslakeid")) %>% 
  dplyr::filter(!is.na(shape_class)) %>%
  dplyr::arrange(desc(factor(shape_class, 
                        levels = c("neither", "concave", "convex"))))

@@

gg_slope_shape <- ggplot(data = test) + 
  geom_point(aes(slope_mean, inlake_slope.y, 
                             color = shape_class), size = 1) +
  geom_smooth(aes(slope_mean, inlake_slope.y, 
                             color = shape_class), data = dplyr::filter(test, shape_class != "neither"), 
              method = "lm", se = FALSE) +
  geom_segment(aes(x = 0, xend = 0.35, y = 0, yend = 0.35), 
               linetype = "dashed", size = 0.4) +
  # geom_abline(slope = 1, intercept = 0) +
  xlab("Nearshore slope (mean)") +
  ylab("In-lake slope") +
  labs(color = "")

gg_slope_reservoir <- ggplot(data = dplyr::filter(test, !is.na(reservoir_class))) +
  geom_point(aes(slope_mean, inlake_slope.y, 
                                 color = reservoir_class), size = 1) +
  geom_smooth(aes(slope_mean, inlake_slope.y, 
                                 color = reservoir_class), 
              method = "lm", se = FALSE) +
  geom_segment(aes(x = 0, xend = 0.35, y = 0, yend = 0.35), 
               linetype = "dashed", size = 0.4) +
  xlab("Nearshore slope (mean)") +
  ylab("In-lake slope") +
  labs(color = "")
```

@@

# slope data
bp <- readRDS("../data/00_bathy_depth/bathy_pnts.rds")
dt <- read.csv("../data/00_geometry/nearshore.csv", stringsAsFactors = FALSE) %>%
  mutate(llid = as.character(llid))
dt <- left_join(dt, st_drop_geometry(bp), 
                by = "llid") %>%
  dplyr::filter(!is.na(inlake_slope_mean)) %>%
  dplyr::filter(inlake_slope.y < 0.4)
lgus_pred <- read.csv("../data/lagosus_depth_predictors.csv", stringsAsFactors = FALSE) %>%
  mutate(lagoslakeid = as.character(lagoslakeid))
test <- left_join(dt, lgus_pred, by = c("llid" = "lagoslakeid")) %>% 
  dplyr::filter(!is.na(shape_class)) %>%
  dplyr::arrange(desc(factor(shape_class, 
                        levels = c("neither", "concave", "convex"))))

gg_slope_base <- ggplot(data = test) + 
  geom_point(aes(slope_mean, inlake_slope.y), size = 1) +
  geom_smooth(aes(slope_mean, inlake_slope.y), method = "lm", se = FALSE) +
  geom_rect(xmin = 0.28, xmax = 0.42, ymin = 0, ymax = 0.05,
  alpha = 1, fill = "white") +
  annotate("text", x = 0.34, y = 0.03, label = "R^2 == 0.17",
                                parse = TRUE, size = text_size) +
  annotate("text", x = 0.34, y = 0.01, 
           label = get_formula(
             lm(inlake_slope.y ~ slope_mean, data = test)), 
           size = text_size) +
  # geom_abline(slope = 1, intercept = 0) +
  xlab("Nearshore slope (mean)") +
  ylab("In-lake slope") +
  labs(color = "") +
  theme_opts

@@

nearshore     <- read.csv("../data/00_geometry/nearshore.csv",
                      stringsAsFactors = FALSE) %>%
                  dplyr::select(-inlake_slope)
dt_raw        <- read.csv("../data/lagosus_depth_predictors.csv", 
                      stringsAsFactors = FALSE) %>%
                  left_join(nearshore, b = c("lagoslakeid" = "llid")) %>% 
                  dplyr::filter(shape_class %in% c("concave", "convex")) %>% 
                  dplyr::filter(inlake_slope < 0.6)

@@

gg_area <- ggplot(data = dt) +
   geom_point(aes(x = slope_mean, y = inlake_slope, color = x_bins)) +
   geom_smooth(aes(slope_mean, inlake_slope, color = x_bins),
               method = "lm", se = FALSE) + 
  facet_wrap(~shape_class) +
    ggtitle("Area (ha)")

@@

gg_depth <- ggplot(data = dt) +
   geom_point(aes(x = slope_mean, y = inlake_slope, color = x_bins)) +
   geom_smooth(aes(slope_mean, inlake_slope, color = x_bins),
               method = "lm", se = FALSE) +
    facet_wrap(~shape_class) +
    ggtitle("Max Depth (m)")

>>>>

```{r slope-compare, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.height=3.5}
# compare inlake slope, buffer slope, and online slope
# setwd("figures") 
bp <- readRDS("../data/00_bathy_depth/bathy_pnts.rds")
dt <- read.csv("../data/00_geometry/nearshore.csv", stringsAsFactors = FALSE) %>%
  mutate(llid = as.character(llid)) %>%
  dplyr::select(-contains("inlake_slope"))
dt <- left_join(dt, st_drop_geometry(bp), 
                by = "llid") %>%
  dplyr::filter(!is.na(inlake_slope_mean))

slope_dist_str <- "online_mean|slope_pnt|inlake_slope|dist_v|dist_d|dists_|slope_mean"
# jsta::jheatmap
jheatmap(
  dplyr::select(dt, 
   matches(slope_dist_str))
   # maxdepth,
   # meandepth)
  )

GGally::ggpairs(dplyr::select(dt, 
                inlake_slope_mean, inlake_slope_pnt,
                nearshore_slope_online_mean,
                nearshore_slope_mean, 
                # dist_deepest, 
                # dist_viscenter, 
                maxdepth), 
                progress = FALSE, 
                ) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1), 
          strip.text.y = element_text(angle = 0))

@@

# dt[22,]
# which(dt$inlake_slope.y > 0.4)
# which(dt$inlake_slope.x > 0.4)
dt <- dplyr::filter(dt, inlake_slope_pnt < 0.4)

cowplot::plot_grid(
  gg_scatter_slope("nearshore_slope_mean", "maxdepth", dt),
  gg_scatter_slope("inlake_slope_pnt", "maxdepth", dt), 
  gg_scatter_slope("nearshore_slope_mean", "inlake_slope_pnt", dt)
  )
```

```{r buffer-vs-inlake_slope, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide', fig.width=4}
lgus_pred <- read.csv("../data/lagosus_depth_predictors.csv", stringsAsFactors = FALSE) %>%
  mutate(lagoslakeid = as.character(lagoslakeid)) %>%
  dplyr::select(-matches("inlake_slope"), 
                -matches("dist_|dists_"))
test <- left_join(dt, lgus_pred, by = c("llid" = "lagoslakeid")) %>% 
  dplyr::filter(!is.na(shape_class)) %>%
  dplyr::arrange(desc(factor(shape_class, 
                        levels = c("neither", "concave", "convex"))))

@@

gg_slope_shape <- ggplot(data = test) + 
  geom_point(aes(nearshore_slope_mean, inlake_slope_pnt, 
                             color = shape_class), size = 1) +
  geom_smooth(aes(nearshore_slope_mean, inlake_slope_pnt, 
                             color = shape_class), data = dplyr::filter(test, shape_class != "neither"), 
              method = "lm", se = FALSE) +
  geom_segment(aes(x = 0, xend = 0.35, y = 0, yend = 0.35), 
               linetype = "dashed", size = 0.4) +
  # geom_abline(slope = 1, intercept = 0) +
  xlab("Nearshore slope (mean)") +
  ylab("In-lake slope") +
  labs(color = "")

gg_slope_reservoir <- ggplot(data = dplyr::filter(test, !is.na(reservoir_class))) +
  geom_point(aes(nearshore_slope_mean, inlake_slope_pnt, 
                                 color = reservoir_class), size = 1) +
  geom_smooth(aes(nearshore_slope_mean, inlake_slope_pnt, 
                                 color = reservoir_class), 
              method = "lm", se = FALSE) +
  geom_segment(aes(x = 0, xend = 0.35, y = 0, yend = 0.35), 
               linetype = "dashed", size = 0.4) +
  xlab("Nearshore slope (mean)") +
  ylab("In-lake slope") +
  labs(color = "")
```

@@

# slope data
bp <- readRDS("../data/00_bathy_depth/bathy_pnts.rds")
dt <- read.csv("../data/00_geometry/nearshore.csv", stringsAsFactors = FALSE) %>%
  mutate(llid = as.character(llid)) %>%
  dplyr::select(-matches("inlake_slope")) # prevent name conflict with bp
dt <- left_join(dt, st_drop_geometry(bp), 
                by = "llid") %>%
  dplyr::filter(!is.na(inlake_slope_mean)) %>%
  dplyr::filter(inlake_slope_pnt < 0.4)
lgus_pred <- read.csv("../data/lagosus_depth_predictors.csv", stringsAsFactors = FALSE) %>%
  mutate(lagoslakeid = as.character(lagoslakeid)) %>%
  dplyr::select(-matches("inlake_slope"))
test <- left_join(dt, lgus_pred, by = c("llid" = "lagoslakeid")) %>% 
  dplyr::filter(!is.na(shape_class)) %>%
  dplyr::arrange(desc(factor(shape_class, 
                        levels = c("neither", "concave", "convex"))))

gg_slope_base <- ggplot(data = test) + 
  geom_point(aes(nearshore_slope_mean, inlake_slope_pnt), size = 1) +
  geom_smooth(aes(nearshore_slope_mean, inlake_slope_pnt), method = "lm", se = FALSE) +
  geom_rect(xmin = 0.28, xmax = 0.42, ymin = 0, ymax = 0.05,
  alpha = 1, fill = "white") +
  annotate("text", x = 0.34, y = 0.03, label = "R^2 == 0.17",
                                parse = TRUE, size = text_size) +
  annotate("text", x = 0.34, y = 0.01, 
           label = get_formula(
             lm(inlake_slope_pnt ~ nearshore_slope_mean, data = test)), 
           size = text_size) +
  # geom_abline(slope = 1, intercept = 0) +
  xlab("Nearshore slope (mean)") +
  ylab("In-lake slope") +
  labs(color = "") +
  theme_opts

@@

nearshore     <- read.csv("../data/00_geometry/nearshore.csv",
                      stringsAsFactors = FALSE) %>%
                  dplyr::select(-matches("inlake_slope")) 
dt_raw        <- read.csv("../data/lagosus_depth_predictors.csv", 
                      stringsAsFactors = FALSE) %>%
                  left_join(nearshore, b = c("lagoslakeid" = "llid")) %>% 
                  dplyr::filter(shape_class %in% c("concave", "convex")) %>% 
                  dplyr::filter(inlake_slope_pnt < 0.6)

@@

gg_area <- ggplot(data = dt) +
   geom_point(aes(x = nearshore_slope_mean, y = inlake_slope_pnt, color = x_bins)) +
   geom_smooth(aes(nearshore_slope_mean, inlake_slope_pnt, color = x_bins),
               method = "lm", se = FALSE) + 
  facet_wrap(~shape_class) +
    ggtitle("Area (ha)")

@@

gg_depth <- ggplot(data = dt) +
   geom_point(aes(x = nearshore_slope_mean, y = inlake_slope_pnt, color = x_bins)) +
   geom_smooth(aes(nearshore_slope_mean, inlake_slope_pnt, color = x_bins),
               method = "lm", se = FALSE) +
    facet_wrap(~shape_class) +
    ggtitle("Max Depth (m)")

### 2020-11-23 tables/01_predictors.Rmd
propagate bathy metric name changes, calculate mape model metric, closes #17

# setwd("tables")
nearshore <- read.csv("../data/00_geometry/nearshore.csv", 
                      stringsAsFactors = FALSE) %>%
  dplyr::select(llid, buffer_slope = slope_mean)
dt_all <- read.csv("../data/lagosus_depth_predictors.csv", 
               stringsAsFactors = FALSE) %>%
  distinct(lagoslakeid, .keep_all = TRUE)
dt_bathy <- dt_all %>%
  right_join(nearshore, by = c("lagoslakeid" = "llid")) %>%
  dplyr::select(-contains("lagoslakeid"), 
                -contains("lake_nhd"), 
                -contains("lake_huc"))

@@

# compute stats for bathymetry lakes
dt_bathy <- compute_stats(dt_bathy)
dt_bathy <- data.frame(dt_bathy)
# compute stats for all of lagosus
dt_all <- data.frame(compute_stats(dt_all))
dt_all <- dplyr::filter(dt_all, variable %in% dt_bathy$variable) 
# interleave
dt_combined <- bind_cols(lapply(names(dt_bathy)[2:5], 
                           function(i) paste0(dt_bathy[,i], 
                                              " (", dt_all[,i], ")"))) %>%
  setNames(names(dt_bathy)[2:5])

@@

```{r predictors_table}
# setwd("figures")
dt <- dt_combined %>%
  dplyr::filter(str_detect(variable, 
                           "^lake_|^ws_lake_arearatio|^dist_|_slope")) %>%
  dplyr::filter(!str_detect(variable,
                      "totalarea|islandperimeter|between|buffer100m|lake_meandepth_m")) # %>%
  # mutate(n = dt_bathy$n[which(dt_bathy$variable == "inlake_slope")])

# remove sample size reporting for bathy specific variables
dt[
  dt$variable %in% c("dist_deepest", "dist_viscenter", 
                     "inlake_slope", "buffer_slope"), 2:5] <- 
  dt[
  dt$variable %in% c("dist_deepest", "dist_viscenter", 
                     "inlake_slope", "buffer_slope"), 2:5] %>%
apply(2, function(x) stringr::str_replace(x, "\\(([^\\)]+)\\)", "(-)"))

@@

# bold Eq 2 predictor variables
dt_md <- dt
dt_md$Variable[!stringr::str_detect(dt_md$Variable, "slope|Distance|depth")] <- 
  paste0("**", 
         dt_md$Variable[!stringr::str_detect(dt_md$Variable, "slope|Distance|depth")],
         "**")

@@

# bold Eq 2 predictor variables
dt$Variable[!stringr::str_detect(dt$Variable, "slope|Distance|depth")] <- 
  paste0("\\textbf{", 
         dt$Variable[!stringr::str_detect(dt$Variable, "slope|Distance|depth")],
         "}")

>>>>

# setwd("tables")
nearshore <- read.csv("../data/00_geometry/nearshore.csv", 
                      stringsAsFactors = FALSE) %>%
  dplyr::select(llid, contains(c("slope_mean", "slope_online_mean", 
                                 "slopes_mean", "slopes_online_mean")))
dt_all <- read.csv("../data/lagosus_depth_predictors.csv", 
               stringsAsFactors = FALSE) %>%
  distinct(lagoslakeid, .keep_all = TRUE) %>%
  dplyr::select(-inlake_slope_mean)
dt_bathy <- dt_all %>%
  right_join(nearshore, by = c("lagoslakeid" = "llid")) %>%
  dplyr::select(-contains("lagoslakeid"), 
                -contains("lake_nhd"), 
                -contains("lake_huc"))

@@

# compute stats for bathymetry lakes
dt_bathy <- compute_stats(dt_bathy)
dt_bathy <- data.frame(dt_bathy)
# compute stats for all of lagosus
dt_all <- data.frame(compute_stats(dt_all))
dt_all <- dplyr::filter(dt_all, variable %in% dt_bathy$variable) 
# interleave
dt_combined <- suppressMessages(
  bind_cols(lapply(names(dt_bathy)[2:5], 
                           function(i) paste0(dt_bathy[,i], 
                                              " (", dt_all[,i], ")")))
  ) %>%
  setNames(names(dt_bathy)[2:5])

@@

```{r predictors_table}
# setwd("figures")
(slope_distance_vars <- unique(dt_combined$variable)[
  stringr::str_detect(unique(dt_combined$variable), 
                     "dist_|e_slope_|dists_|e_slopes")])
(exclude_vars <- unique(dt_combined$variable)[
  stringr::str_detect(unique(dt_combined$variable), 
                     "totalarea|islandperimeter|between|buffer100m|lake_meandepth_m")])

dt <- dt_combined %>%
  dplyr::filter(str_detect(variable, 
                c(paste0(slope_distance_vars, collapse = "|"),
                  "^lake_|^ws_lake_arearatio"))) %>%
  dplyr::filter(!str_detect(variable,
                            paste0(exclude_vars, collapse = "|")))
                       # %>%
  # mutate(n = dt_bathy$n[which(dt_bathy$variable == "inlake_slope")])

# remove sample size reporting for bathy specific variables
dt[
  dt$variable %in% slope_distance_vars, 2:5] <- 
  dt[
  dt$variable %in% slope_distance_vars, 2:5] %>%
apply(2, function(x) stringr::str_replace(x, "\\(([^\\)]+)\\)", "(-)"))

@@

# bold Eq 2 predictor variables
dt_md <- dt
dt_md$Variable[!stringr::str_detect(dt_md$Variable, "slope|Distance|depth") & !is.na(dt_md$Variable)] <- 
  paste0("**", 
         dt_md$Variable[!stringr::str_detect(dt_md$Variable, "slope|Distance|depth") & !is.na(dt_md$Variable)],
         "**")

@@

# bold Eq 2 predictor variables
dt$Variable[!stringr::str_detect(dt$Variable, "slope|Distance|depth") & !is.na(dt_md$Variable)] <- 
  paste0("\\textbf{", 
         dt$Variable[!stringr::str_detect(dt$Variable, "slope|Distance|depth") & !is.na(dt_md$Variable)],
         "}")

### 2020-11-23 tables/02_model_metrics.Rmd
propagate bathy metric name changes, calculate mape model metric, closes #17

dt_metrics <- readRDS("../data/01_depth_model/depth_grid_metrics.rds") %>%
  mutate_if(is.numeric, function(x) as.character(round(x, 2))) %>%
  mutate(rsq = as.numeric(rsq)) %>%
  mutate(rsq = case_when(
    as.character(model) == "maxdepth_true_true" ~ NA_real_,
    TRUE ~ rsq)) %>%
  mutate(rmse = round(as.numeric(rmse), 1)) %>%
  mutate(rmse = case_when(
    as.character(model) == "maxdepth_true_true" ~ NA_character_,
    TRUE ~ paste0(rmse, " m"))) %>%
  mutate(model = gsub("maxdepth_", "", model)) %>%
  mutate(model = gsub("false", "proxy", model)) %>%
  tidyr::separate(model, c("slope", "distance")) %>%
  dplyr::select(slope, distance, everything(), -.estimator) %>%
  dplyr::select(-mae)

>>>>

dt_metrics <- readRDS("../data/01_depth_model/depth_grid_metrics.rds") %>%
  mutate_if(is.numeric, function(x) as.character(round(x, 2))) %>%
  mutate(rsq = as.numeric(rsq)) %>%
  mutate(rsq = case_when(
    as.character(model) == "maxdepth_true_true" ~ NA_real_,
    TRUE ~ rsq)) %>%
  mutate(rmse = round(as.numeric(rmse), 1)) %>%
  mutate(rmse = case_when(
    as.character(model) == "maxdepth_true_true" ~ NA_character_,
    TRUE ~ paste0(rmse, " m"))) %>%
  mutate(model = gsub("maxdepth_", "", model)) %>%
  mutate(model = gsub("false", "proxy", model)) %>%
  tidyr::separate(model, c("slope", "distance")) %>%
  dplyr::select(slope, distance, everything(), -.estimator)
